Working title: Highly resolved spatiotemporal simulations for exploring mixed fishery dynamics

### 2019-09-17 write_up/mixfishsim_ecolmod.tex
Minor tweak

%%
To understand how data resolution impacts inference on mixed fisheries
interactions, we developed a highly resolved spatiotemporal discrete-event
simulation model incorporating: i) delay-difference population dynamics, ii)
population movement using Gaussian Random Fields to simulate patchy,
heterogeneously distributed populations, and iii) fishery dynamics for multiple
fleet characteristics based on species targeting under an explore-exploit
strategy. This is implemented via a mix of correlated random walk movement (for
exploration) and learned behaviour (for exploitation) phases of the fisheries.
\\ 
%%
Fifty years of sub-daily fishing was simulated and used to draw inference on
the underlying community structures. We compared inferences based on:
commercial catch, a simulated fixed-site sampling survey design and the true
underlying community. We i) establish the potential limitations of
fishery-dependent data in providing a robust picture of spatiotemporal
distributions; and then ii) simulated an area closure based on areas defined
from the different data sources at a range of temporal and spatial resolutions.
\\
%%
Our framework allows users to explore the assumptions in modelling
observational data and evaluate the underlying dynamics of such approaches at a
fine spatial and temporal scale. We conclude from our example framework
application that commercial data, while containing bias, provide a useful tool
for managing catches in mixed fisheries if applied at the correct
spatiotemporal scale. [NEED A LITTLE MORE DETAIL OF THE RESULTS HERE - 2 MORE
SENTENCES ON KEY TAKE-HOME RESULTS]\\
%%
\end{abstract}

@@

The final combined process results in a population structure and movement
pattern unique to each species, with population movement occurring on a weekly
basis. The decision to model population movement on a weekly timescale was to
reflect that fish tend to aggregate in species specific locations that have
been observed to last around one to two weeks \citep{Poos2007}. Therefore this
process approximated the demographic shifts in fish populations throughout a
year with seasonal spawning patterns (Figure \ref{S-fig:5}).

@@

Fleet dynamics can be broadly categorised into three components: fleet
targeting - that determined the fleet catch efficiency and preference towards a
particular species; trip-level decisions, that determines the initial location
to be fished at the beginning of a trip; and within-trip decisions, that
determines movement from one fishing spot to another within a trip. An
explore-exploit type strategy was implemented in the model that combined these
three components for individual vessels to maximise their catch from an unknown
resource distribution \citep{Bailey2018}. The decision to use an individual
based model for fishing vessels was taken because fishers are heterogeneous in
their location choice behaviour due to different objectives, risk preference
and targeting preference \citep{VanPutten2012a, Boonstra2015}. Therefore fleet
dynamics are emergent from individual dynamics rather than pre-defined group
dynamics.

@@

Each fleet of \textit{n} vessels was characterised by both a general
efficiency, $Q_{fl}$, and a population specific efficiency, ${Q_{fl, p}}$ which
are each bound by [0,1]. The product of these parameters [$Q_{fl} \cdot
Q_{fl, p}$] affects the overall catch rates for the fleet and the preferential
targeting of one population over another. This, in combination with the
parameter choice for the step-function defined below (as well as some
randomness from the exploratory fishing process) determined the preference of
fishing locations for the fleet.

@@

For our implementation of a random walk directional change is based on a
negatively correlated circular distribution where a favourable fishing ground
is likely to be ``fished back over" by the vessel returning in the direction it
came from.  The step length (i.e. the distance travelled from the current to
the next fishing location) is determined by relating recent
fishing success, measured as the summed value of fish caught (revenue, $Rev$);
\begin{equation}
Rev_{c,d} = \sum_{p=1}^{P} L_{c,d,p} \cdot Pr_{p} 
\end{equation}
where $L_{c,d,p}$ is landings of a population $p$, and $Pr_{p}$ price of a
population. All species prices were kept the same across fleets and seasons.
Here, when fishing is successful vessels remain in a similar location and
continue to exploit the local fishing grounds. When unsuccessful, they move
some distance away from the current fishing location.  The movement distance
retains some degree of stochasticity, that can be controlled separately, but is
determined by the relationship: 
\begin{equation}
	Le = \mathrm{e}^{log(\beta_{1}) + log(\beta_{2}) -
		\left(log\left(\frac{\beta_{1}}{\beta_{3}}\right)\right) \cdot Rev
} 
\end{equation}
Where $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$ are parameters determining the
shape of the step function in its relation to revenue, so that, a step from
(x1,y1) to (x2, y2) is defined by:
\begin{equation}
	\begin{split}
 (x2, y2) =  & x1 + Le \cdot \cos \left(\frac{\pi \cdot Br}{180}\right), \\
             & y1 + Le \cdot \sin \left(\frac{\pi \cdot Br}{180}\right) \\	
 with  \hspace{0.5cm}     & Br_{t-1} < 180, Br_{t} = 180 + \sim vm[(0,360), k] \\
 			  & Br_{t-1} > 180, Br_{t} = 180 - \sim vm[(0,360), k] \\
	\end{split}
\end{equation}
where $Le$ is the step length, $k$ the concentration parameter from the von
Mises distribution that we correlate with the revenue so that $k = (Rev + 1 /
RefRev) \cdot max_{k}$, where $max_{k}$ is the maximum concentration value,
$k$, and $RefRev$ is parametrised as for $\beta_{3}$ in the step length
function.

@@

To answer this question we compare different spatial and temporal aggregations
of the true population distributions to:
\begin{enumerate}[label=\alph*)]
	\item \textbf{fisheries-independent data:} the inferred population from
		a fixed-site sampling survey design as commonly used for
		fisheries monitoring purposes;
	\item \textbf{fisheries-dependent data:} the inferred population from
		our fleet model that includes fishery-induced sampling
		dynamics.
\end{enumerate}

@@

\begin{itemize}
	\item \textbf{data types:} commercial logbook data, survey data and
		true population,
	\item \textbf{temporal resolutions:} weekly, monthly and yearly
		closures,
	\item \textbf{spatial resolutions:} 1 x 1 grid, 5 x 5 grid, 10 x 10
		grid and 20 x 20 grid,
	\item \textbf{closure basis:} highest 5 \%  of catch rates for the
		protected species
\end{itemize}

@@

Different perceptions of the proportion of each stock in an area are seen when
we aggregate the data over a smaller geographical region at different
timescales, with weekly (top), monthly (middle) and yearly (bottom) catch
compositions from across an aggregated 20 x 20 area (Figure \ref{fig:2}). In
the true population, the monthly aggregation captures the major patterns of
composition seen in the weekly data with the percentage of different species in
the catch having similar mean and standard deviations.  In the weekly data
population 1 = 9.36 (3.99), population 2 = 83.2 (5.60), population 3 = 3.57 (1.23),
population 4 = 3.91 (1.59); in the monthly data population 1 = 9.23 (3.87), population 2
= 83.3 (5.52), population 3 = 3.62 (1.15), population 4 = 3.86 (1.52). While means
were similar some of the variation was lost when aggregated to an annual level;
population 1 = 9.90 (0.173), population 2 = 82.2 (0.308), population 3 = 3.82 (0.119),
population 4 = 4.03 (0.0502).\\

>>>>

%%
To understand how data resolution impacts inference on mixed fisheries
interactions, we developed a highly resolved spatiotemporal discrete-event
simulation model incorporating: i) delay-difference population dynamics, ii)
population movement using Gaussian Random Fields to simulate patchy,
heterogeneously distributed populations, and iii) fishery dynamics for multiple
fleet characteristics based on population targeting under an explore-exploit
strategy. This is implemented via a mix of correlated random walk movement (for
exploration) and learned behaviour (for exploitation) phases of the fisheries.
\\ 
%%
Fifty years of sub-daily fishing was simulated and used to draw inference on
the underlying community structures. We compared inferences based on:
commercial catch, a simulated fixed-site sampling survey design and the true
underlying community. We i) establish the potential limitations of
fishery-dependent data in providing a robust picture of spatiotemporal
distributions; and then ii) simulated an area closure based on areas defined
from the different data sources at a range of temporal and spatial resolutions.
\\
%%
Our framework allows users to explore the assumptions in modelling
observational data and evaluate the underlying dynamics of such approaches at a
fine spatial and temporal scale. We conclude from our example framework
application that commercial data, while containing bias, provide a useful tool
for managing catches in mixed fisheries if applied at the correct
spatiotemporal scale. [NEED A LITTLE MORE DETAIL OF THE RESULTS HERE - 2 MORE
SENTENCES ON KEY TAKE-HOME RESULTS]\\
%%
\end{abstract}

@@

The final combined process results in a population structure and movement
pattern unique to each population, with population movement occurring on a weekly
basis. The decision to model population movement on a weekly timescale was to
reflect that fish tend to aggregate in species specific locations that have
been observed to last around one to two weeks \citep{Poos2007}. Therefore this
process approximated the demographic shifts in fish populations throughout a
year with seasonal spawning patterns (Figure \ref{S-fig:5}).

@@

Fleet dynamics can be broadly categorised into three components: fleet
targeting - that determined the fleet catch efficiency and preference towards a
particular population; trip-level decisions, that determines the initial location
to be fished at the beginning of a trip; and within-trip decisions, that
determines movement from one fishing spot to another within a trip. An
explore-exploit type strategy was implemented in the model that combined these
three components for individual vessels to maximise their catch from an unknown
resource distribution \citep{Bailey2018}. The decision to use an individual
based model for fishing vessels was taken because fishers are heterogeneous in
their location choice behaviour due to different objectives, risk preference
and targeting preference \citep{VanPutten2012a, Boonstra2015}. Therefore fleet
dynamics are emergent from individual dynamics rather than pre-defined group
dynamics.

@@

Each fleet of \textit{n} vessels was characterised by both a general
efficiency, $Q_{fl}$, and a population specific efficiency, ${Q_{fl, p}}$ which
are each bound by [0,1]. The product of these parameters [$Q_{fl} \cdot
Q_{fl, p}$] affects the overall catch rates for the fleet and the preferential
targeting of one species over another. This, in combination with the
parameter choice for the step-function defined below (as well as some
randomness from the exploratory fishing process) determined the preference of
fishing locations for the fleet.

@@

For our implementation of a random walk directional change is based on a
negatively correlated circular distribution where a favourable fishing ground
is likely to be ``fished back over" by the vessel returning in the direction it
came from.  The step length (i.e. the distance travelled from the current to
the next fishing location) is determined by relating recent
fishing success, measured as the summed value of fish caught (revenue, $Rev$);
\begin{equation}
Rev_{c,d} = \sum_{p=1}^{P} L_{c,d,p} \cdot Pr_{p} 
\end{equation}
where $L_{c,d,p}$ is landings of a population $p$, and $Pr_{p}$ price of a
population. All population prices were kept the same across fleets and seasons.
Here, when fishing is successful vessels remain in a similar location and
continue to exploit the local fishing grounds. When unsuccessful, they move
some distance away from the current fishing location.  The movement distance
retains some degree of stochasticity, that can be controlled separately, but is
determined by the relationship: 
\begin{equation}
	Le = \mathrm{e}^{log(\beta_{1}) + log(\beta_{2}) -
		\left(log\left(\frac{\beta_{1}}{\beta_{3}}\right)\right) \cdot Rev
} 
\end{equation}
Where $\beta_{1}$, $\beta_{2}$ and $\beta_{3}$ are parameters determining the
shape of the step function in its relation to revenue, so that, a step from
(x1,y1) to (x2, y2) is defined by:
\begin{equation}
	\begin{split}
 (x2, y2) =  & x1 + Le \cdot \cos \left(\frac{\pi \cdot Br}{180}\right), \\
             & y1 + Le \cdot \sin \left(\frac{\pi \cdot Br}{180}\right) \\	
 with  \hspace{0.5cm}     & Br_{t-1} < 180, Br_{t} = 180 + \sim vm[(0,360), k] \\
 			  & Br_{t-1} > 180, Br_{t} = 180 - \sim vm[(0,360), k] \\
	\end{split}
\end{equation}
where $Le$ is the step length, $k$ the concentration parameter from the von
Mises distribution that we correlate with the revenue so that $k = (Rev + 1 /
RefRev) \cdot max_{k}$, where $max_{k}$ is the maximum concentration value,
$k$, and $RefRev$ is parametrised as for $\beta_{3}$ in the step length
function.

@@

To answer this question we compare different spatial and temporal aggregations
of the true population distributions to:
\begin{enumerate}[label=\alph*)]
	\item \textbf{fisheries-independent data:} the inferred population
		density from a fixed-site sampling survey design as commonly
		used for
		fisheries monitoring purposes;
	\item \textbf{fisheries-dependent data:} the inferred population
		density from our fleet model that includes fishery-induced
		sampling dynamics.
\end{enumerate}

@@

\begin{itemize}
	\item \textbf{data types:} commercial logbook data, survey data and
		true population,
	\item \textbf{temporal resolutions:} weekly, monthly and yearly
		closures,
	\item \textbf{spatial resolutions:} 1 x 1 grid, 5 x 5 grid, 10 x 10
		grid and 20 x 20 grid,
	\item \textbf{closure basis:} highest 5 \%  of catch rates for the
		protected population
\end{itemize}

@@

Different perceptions of the proportion of each stock in an area are seen when
we aggregate the data over a smaller geographical region at different
timescales, with weekly (top), monthly (middle) and yearly (bottom) catch
compositions from across an aggregated 20 x 20 area (Figure \ref{fig:2}). In
the true population, the monthly aggregation captures the major patterns of
composition seen in the weekly data with the percentage of different
populations in the catch having similar mean and standard deviations.  In the
weekly data population 1 = 9.36 (3.99), population 2 = 83.2 (5.60), population
3 = 3.57 (1.23), population 4 = 3.91 (1.59); in the monthly data population 1 =
9.23 (3.87), population 2 = 83.3 (5.52), population 3 = 3.62 (1.15), population
4 = 3.86 (1.52). While means were similar some of the variation was lost when
aggregated to an annual level; population 1 = 9.90 (0.173), population 2 = 82.2
(0.308), population 3 = 3.82 (0.119), population 4 = 4.03 (0.0502).\\
